XXXXX

### 2014-10-07 paper/paper1.tex
DC 2014 tutorial
tutorial DC 2014
ideas of Kai

@@

\begin{ex}
:SSN 
    a rdfs:Datatype ;
    owl:equivalentClass [
        a rdfs:Datatype ;
        owl:onDatatype xsd:string ;
        owl:withRestrictions ( 
            [ xsd:pattern "[0-9]{3}-[0-9]{2}-[0-9]{4}" ] ) ] .
\end{ex}

@@

\begin{ex}
:IimBernersLee
    :hasSSN "123-45-6789"^^:SSN .
\end{ex}

@@

\begin{ex}
:IimBernersLee
    :hasSSN "123456789"^^:SSN .
\end{ex}

@@

@@

\begin{ex}
:NumberPlayersPerWorldCupTeam
    a rdfs:Datatype ;
    owl:equivalentClass [
        a rdfs:Datatype ;
        owl:onDatatype xsd:nonNegativeInteger ;
        owl:withRestrictions ( 
            [ xsd:minInclusive "1"^^xsd:nonNegativeInteger ]
            [ xsd:maxInclusive "23"^^xsd:nonNegativeInteger ] ) ] .
\end{ex}

@@

\begin{itemize}
	\item instances of specific class must not have some object property
  \item OWL 2 DL: ObjectComplementOf ( ObjectSomeValuesFrom ( ObjectPropertyExpression owl:Thing ) )
	\item covering approaches: ShEx, SPIN
\end{itemize}

>>>>

Ideas of Kai:
RDF validation using OWL 2 QL reasoning by SPARQL query expansion
validation with mappings to SPARQL
if using OWL 2 DL as constraint language or using constraints equivalent to OWL 2 QL you can use reasoning 
reasoning can also be executed SPARQL query expansion.
reasoning not executing using reasoner
limitations of this approach
expressivity of SPARQL? (Erman) what can not be done with SPARQL?

@@

\begin{ex}
:SSN 
    a rdfs:Datatype ;
    owl:equivalentClass [
        a rdfs:Datatype ;
        owl:onDatatype xsd:string ;
        owl:withRestrictions ( 
            [ xsd:pattern "[0-9]{3}-[0-9]{2}-[0-9]{4}" ] ) ] .
:hasSSN rdfs:range :SSN .
\end{ex}

@@

\begin{ex}
:TimBernersLee
    :hasSSN "123-45-6789"^^:SSN .
\end{ex}

@@

\begin{ex}
:TimBernersLee
    :hasSSN "123456789"^^:SSN .
\end{ex}

@@

covering approaches: DQTP, OWL 2 DL, SPARQL, SPIN

@@

\begin{ex}
:NumberPlayersPerWorldCupTeam
    a rdfs:Datatype ;
    owl:equivalentClass [
        a rdfs:Datatype ;
        owl:onDatatype xsd:nonNegativeInteger ;
        owl:withRestrictions ( 
            [ xsd:minInclusive "1"^^xsd:nonNegativeInteger ]
            [ xsd:maxInclusive "23"^^xsd:nonNegativeInteger ] ) ] .
:position rdfs:range :NumberPlayersPerWorldCupTeam .
\end{ex}

@@

\begin{itemize}
	\item instances of specific class must not have some object property
  \item OWL 2 DL: ObjectComplementOf ( ObjectSomeValuesFrom ( ObjectPropertyExpression owl:Thing ) )
	\item covering approaches: ShEx, SPIN, SPARQL
\end{itemize}

### 2014-10-07 paper/paper3.tex
ideas of Kai

>>>>

\section{XX}

ontology describing constraint languages
interoperability between constraint languages
express constraint in CL A and translate constraint in CL B
provide SPARQL validation (SPIN mapping) for each general C

\section{constraint language ontology}

\section{examples}

\subsection{min cardinality}

OWL 2:
\begin{ex}
foaf:Person
    a owl:Restriction ;
		owl:minQualifiedCardinality "2"^^xsd:nonNegativeInteger ;
		owl:onProperty :hasName ;
		owl:onClass xsd:string .
\end{ex}

general constraint: min cardinality
constraint (OWL2): data min cardinality

translation to SPARQL:
\begin{ex}

\end{ex}

\subsection{object property range}

DSCL constraint (OWL 2):
\begin{ex}
:hasDog rdfs:range :Dog . 
\end{ex}

DSCL constraint (DSP):
\begin{ex}
:hasDogRange
        a dsp:DescriptionTemplate ; 
        dsp:resourceClass owl:Thing ; 
        dsp:statementTemplate [
            a dsp:NonLiteralStatementTemplate ;
            dsp:property :hasDog ; 
            dsp:nonLiteralConstraint [ 
                a dsp:NonLiteralConstraint ;
                dsp:valueClass :Dog ] ] .
\end{ex}

general constraint:
\begin{ex}
clo:objectPropertRange
    a clo:GeneralConstraint ;
		clo:requirement clo:R-28-OBJECT-PROPERTY-RANGE ;
		clo:DSCLConstraints ( clo:objectPropertRange_OWL2 clo:objectPropertRange_DSP ) ;
		clo:objectProperty :hasDog ;
		clo:range ( :Dog ) ;
\end{ex}

valid data:
\begin{ex}
:Peter 
    :hasDog :Brian .
:Brian 
    a :Dog ;     
    a owl:Thing . 
\end{ex}

invalid data:
\begin{ex}
:Peter 
    :hasDog :Brian .
:Brian 
#   a :Dog ; # commented --> constraint violation
    a owl:Thing . 
\end{ex}

general constraint: object property range
constraint (OWL2): object property range

general C is translated to SPARQL for automatic validation

for each DSCL constraint you have to define the binding to the general constraint.

binding (DSCL constraint (OWL2) - general constraint):
\begin{ex}
# OWL 2 constraint
# -----
?OPE rdfs:range ?CE .

clo:objectProperty => ?OPE
clo:range => ?CE 
\end{ex}

binding (DSCL constraint (OWL2) - general constraint):
\begin{ex}
# DSP constraint
# -----
:hasDogRange
        a dsp:DescriptionTemplate ; 
        dsp:resourceClass owl:Thing ; 
        dsp:statementTemplate [
            a dsp:NonLiteralStatementTemplate ;
            dsp:property :hasDog ; 
            dsp:nonLiteralConstraint [ 
                a dsp:NonLiteralConstraint ;
                dsp:valueClass :Dog ] ] .

# SPIN
# -----
?descriptionTemplate a dsp:DescriptionTemplate .
?descriptionTemplate dsp:resourceClass ?resourceClass .
?descriptionTemplate dsp:statementTemplate ?statementTemplate .
?statementTemplate dsp:property ?property .
?statementTemplate dsp:nonLiteralConstraint ?nonLiteralConstraint .
?nonLiteralConstraint dsp:valueClass ?valueClass .				
FILTER EXISTS { ?nonLiteralConstraint dsp:valueClass ?valueClass . }
FILTER NOT EXISTS { ?object a ?valueClass . ?nonLiteralConstraint dsp:valueClass ?valueClass . }

# binding (DSCL constraint (OWL2) - general constraint)
# -----
clo:objectProperty => ?OPE
clo:range => ?CE 
\end{ex}

binding using SPIN

SPARQL (validation of general constraint):
\begin{ex}
?x ?clo:objectProperty ?this .
?this rdf:type owl:Thing . 
FILTER NOT EXISTS { ?this a ?clo:range } .
\end{ex}

SPARQL (validation of DSCL constraint (OWL2)):
\begin{ex}
# constraint
# -----
?OPE rdfs:range ?CE .

# data 
# ----- 
?x ?OPE ?this .
?this rdf:type owl:Thing . 
FILTER NOT EXISTS { ?this rdf:type ?CE } .
\end{ex}

SPARQL (validation of DSCL constraint (DSP)):
\begin{ex}
# data 
# -----
?this a ?resourceClass .
?this ?property ?object .
FILTER ( ?property != rdf:type ) .

# constraint
# -----
?descriptionTemplate a dsp:DescriptionTemplate .
?descriptionTemplate dsp:resourceClass ?resourceClass .
?descriptionTemplate dsp:statementTemplate ?statementTemplate .
?statementTemplate dsp:property ?property .
?statementTemplate dsp:nonLiteralConstraint ?nonLiteralConstraint .
?nonLiteralConstraint dsp:valueClass ?valueClass .				
FILTER EXISTS { ?nonLiteralConstraint dsp:valueClass ?valueClass . }
FILTER NOT EXISTS { ?object a ?valueClass . ?nonLiteralConstraint dsp:valueClass ?valueClass . }
\end{ex}

ontology design pattern.org
constraint grammar
constraint elements are the same
DSP uses other design pattern / 
constraint / constraint elements / constraint design patterns / constraint language
define terminology
related to requirements DB / relate requirements to constraints, constraint languages
constraint element: e.g. min cardinality
dependency between requirements / when this requirements is fulfilled then this is also fulfilled (e.g. min card and requ. car)
min car more powerful than req. propery
why ontology? automatic translation from DSP to OWL 
define general constraint language
we do not invent a new constraint languages
requirements in SPARQL umsetzten
only one validation for each requirement
constraints may be composed
